Analysising social factors influence Canadian voters' choices

### 2024-04-02 paper/paper.qmd
update

---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - Yichen ji
thanks: "Code and data are available at: https://github.com/Selinayichenji/Canadian_election_prediction.git."
date: today
date-format: long
abstract: "1) what was done, 2) what was found, and 3) why this matters (all at a high level). ."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r setup}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(arrow)
```

```{r}
#| label: cleaned
#| warning: false
#| message: false
#| results: 'hide'
#| echo: false

analysis_data <- read_parquet("data/analysis_data/analysis_data.parquet")

@@

@@

The survey data was sourced from the "2021 Canadian Election Study" (Stephenson, Harell, Rubenson, & Loewen, 2022), and the census information was obtained from the "General Social Survey, Cycle 31: Families, Public Use Microdata File Documentation and User’s Guide" (Statistics Canada, 2019).

@@

## Sample of cleaned data

## Measurement
The CES 2021, consisting of an online cohort of 20,968 participants from the broader Canadian populace, was facilitated via the Leger Opinion panel. This study was structured in two distinct stages: the campaign period survey (CPS) and the post-election survey (PES). The CPS phase was further divided into three segments – "CPS", "CPS Modules", and "CPS Oversample", which were later consolidated into one dataset. Regarding the PES, all individuals who participated in the CPS were invited back for this subsequent survey after the election. Out of these, 15,069 individuals completed the PES, marking a response rate of 72%.

@@

distribution of age, gender, income, education level, province, children number

# Model

@@

We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

@@

# Results

```{r echo=FALSE}
# Retrieve the Bayesian linear model from the saved .rds file
liberal_model <- readRDS(file = "models/liberal_model.rds")

@@

# Discussion
What is done in this paper? What is something that we learn about the world? What is another thing that we learn about the world? What are some weaknesses of what was done? What is left to learn or how should we proceed in the future?

@@

# Appendix {-}

@@

## Posterior predictive check

>>>>

---
title: "Analysising social factors influence Canadian voters' choices"
subtitle: "Decoding CES 2021 Vote Data by using A Multilevel Logistic Regression Model"
author: 
  - Yichen ji
thanks: "Code and data are available at: https://github.com/Selinayichenji/Canadian_election_prediction.git."
date: today
date-format: long
abstract: "1) what was done, 2) what was found, and 3) why this matters (all at a high level). ."
format: pdf
toc: true
number-sections: true
bibliography: references.bib
link-citations: true
---

```{r setup}
#| include: false
#| warning: false
#| message: false
library(tidyverse)

library(here)
library(dplyr)
library(arrow)
library(ggplot2)
library(kableExtra)
library(bayesplot)
library(rstanarm)

@@

```{r include=FALSE, message=FALSE}
# read the data file and store it in the 'data' variable
analysis_data <- read_parquet(here("data", "analysis_data", "analysis_data.parquet"))
```

@@

The survey data was sourced from the "2021 Canadian Election Study" (Stephenson, Harell, Rubenson, & Loewen, 2022).

@@

\newpage
## Sample of cleaned data

```{r}
#| label: tbl-cleaned
#| tbl-cap: Sample of the Cleaned Data
#| warning: false
#| echo: false

head(analysis_data, 5) |>
  kable(
    col.names = c("age", "vote_liberal", "vote_conservative", "vote_NDP","gender","education_level","family_income","province","children_number"),
    booktabs = TRUE
  )

@@

All detailed tables include summary data of 6 variables are in @sec-A
age, gender, income, education level, province, children number

```{r}
#| label: fig-pro
#| fig-cap:  Population Distribution by Province in CES2021
#| warning: false
#| echo: false

province_plot <- analysis_data %>%
  count(province) %>%
  mutate(percentage = n / sum(n) * 100)

ggplot(province_plot, aes(x = province, y = percentage, fill = province)) +
  geom_bar(stat = "identity",fill = "navyblue") +
  labs(x = "Province", y = "Percentage of Total") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) 
```

```{r}
library(dplyr)
library(ggplot2)

# Define a function to calculate percentages for each category within a variable
calculate_percentage <- function(data, category) {
  data %>%
    count(!!sym(category)) %>%
    mutate(percentage = n / sum(n) * 100) %>%
    rename(Category = !!sym(category)) %>%
    mutate(Variable = category)  # store the variable name
}

# Apply the function to each of the variables you are interested in
category_list = c("age", "gender", "family_income", "education_level", "province", "children_number")
percentages <- lapply(category_list, calculate_percentage, data = analysis_data)

# Combine the list of data frames into one data frame
percentages_df <- bind_rows(percentages)

# Generate the plot
combined_plot <- ggplot(percentages_df, aes(x = Category, y = percentage, fill = Variable)) +
  geom_bar(stat = "identity") +
  facet_wrap(~Variable, scales = "free_x") + # allows each facet to have its own x axis
  labs(x = "Category", y = "Percentage of Total") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_fill_brewer(palette = "Set1")

# Print the combined plot
combined_plot

## Measurement
The CES 2021, consisting of an online cohort of 20,968 participants from the broader Canadian populace, was facilitated via the Leger Opinion panel. This study was structured in two distinct stages: the campaign period survey (CPS) and the post-election survey (PES). The CPS phase was further divided into three segments – "CPS", "CPS Modules", and "CPS Oversample", which were later consolidated into one dataset. Regarding the PES, all individuals who participated in the CPS were invited back for this subsequent survey after the election. Out of these, 15,069 individuals completed the PES, marking a response rate of 72%. Respondents needed to be aged 18 or over and Canadian citizens or permanent residents in order to participate.

# Model {#sec-models}

@@

We run the model in R using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.
The `stan_glm` function from `rstanarm`

@@

# Results {#sec-result}

```{r}
#| label: models
#| warning: false
#| message: false
#| echo: false

liberal_model <- readRDS("../models/liberal_model.rds")
cons_model <- readRDS("../models/conservative_model.rds")
NDP_model <- readRDS("../models/NDP_model.rds")
```

## Model Coefficients

```{r}
#| label: tbl-liberal
#| tbl-cap: Coefficients of social factors influence vote for liberal party
#| warning: false
#| echo: false
library(modelsummary)

modelsummary(
  list(
    "Vote for Liberal Party" = liberal_model 
  ),
  statistic = "mad", 
  booktabs = TRUE 
)

@@

# Discussion {#sec-dis}
What is done in this paper? What is something that we learn about the world? What is another thing that we learn about the world? What are some weaknesses of what was done? What is left to learn or how should we proceed in the future?

@@

# Appendix {#sec-A}

@@

## Posterior predictive check {#sec-posterior}

```{r fig-posterior-plot, echo=FALSE, fig.cap="pp check"}

pp_check(liberal_model) + 
  theme_classic() +
  theme(legend.position = "bottom")

pp_check(cons_model) + 
  theme_classic() +
  theme(legend.position = "bottom")

pp_check(NDP_model) + 
  theme_classic() +
  theme(legend.position = "bottom")
```
